---
title: "Tables in `R` Markdown"
author: "Wallace Campbell"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
 ioslides_presentation:
    widescreen: TRUE
    fig_width: 10
    css: css/format.css
    logo: misc/ace_analytics_logo.png
---

## Standard table methods look good, but there is no flexibility and they take up a lot of space!

```{r,echo=FALSE,message=FALSE,results='asis', warning=FALSE,comment=FALSE}

data(mtcars)
library(knitr)
kable(mtcars[1:5,])

```


## Changing table properties using css file


## Can also use `pander` to create the same table


```{r,echo=FALSE,message=FALSE,results='asis', warning=FALSE,comment=FALSE}

library(pander)
pander(mtcars[1:5,])

```



## Use ReporteRs package

```{r,echo=FALSE,message=FALSE,results='asis', warning=FALSE,comment=FALSE}

# http://davidgohel.github.io/ReporteRs/FlexTable.html

library(ReporteRs)
library(boot)
library(dplyr)








# FlexTable( data = mtcars[1:10,], add.rownames = TRUE, 
#   header.cell.props = cellProperties( background.color = "#00557F" ), 
#   header.text.props = textProperties( color = "white", 
#     font.size = 11, font.weight = "bold" ), 
#   body.text.props = textProperties( font.size = 10 )
# )

cat(as.html(FlexTable( data = mtcars[1:10,], add.rownames = TRUE, 
  header.cell.props = cellProperties( background.color = "#00557F" ), 
  header.text.props = textProperties( color = "white", 
    font.size = 11, font.weight = "bold" ), 
  body.text.props = textProperties( font.size = 10 )
)))



# ```{r, results='asis'}
# library(ReporteRs)
# tabl = FlexTable( mtcars,
#                   header.cell.props = cellProperties( background.color = "#003366" ),
#                   header.text.props = textBold( color = "white" ),
#                   add.rownames = TRUE )
# tabl = setZebraStyle( tabl, odd = "#DDDDDD", even = "#FFFFFF" )
# cat(as.html(tabl))
# ```



# library(ReporteRs)
# # Create a docx object
# doc = docx()
# # add a document title
# doc = addParagraph( doc, "A FlexTable example", stylename = "TitleDoc" )
# # add a section title
# doc = addTitle( doc, "How to turn a data.frame into a Word table", level = 1 )
# # some text
# doc = addParagraph( doc, "We use the mtcars dataset as example.", stylename = "DocDefaults" )
# # add a table
# MyFTable = FlexTable( data = mtcars[1:10, ], add.rownames = TRUE )
# # format body content
# MyFTable[3:4, "cyl"] = cellProperties( background.color = "red" )
# MyFTable["Valiant", ] = cellProperties( background.color = "blue" )
# doc = addFlexTable(doc, MyFTable)
# # write the doc
# writeDoc( doc, file = "exemple.docx" )
# # open the Word doc
# browseURL("exemple.docx")

```



## rtable makes it easier to output in Markdown


```{r,echo=FALSE,message=FALSE,results='asis', warning=FALSE,comment=FALSE}
# install.packages("rtable")
library(rtable)
# cat(as.html(vanilla.table(mtcars[1:10,])))
knit_print(FlexTable( data = mtcars[1:10,], add.rownames = TRUE, 
  header.cell.props = cellProperties( background.color = "#00557F" ), 
  header.text.props = textProperties( color = "white", 
    font.size = 11, font.weight = "bold" ), 
  body.text.props = textProperties( font.size = 10 )
))
```


## Short version

```{r,echo=FALSE,message=FALSE,results='asis', warning=FALSE,comment=FALSE}
cat(as.html(vanilla.table(mtcars[1:10,])))
```



## Zebra tables

```{r,echo=FALSE,message=FALSE,results='asis', warning=FALSE,comment=FALSE}

cat(as.html(setZebraStyle( vanilla.table(mtcars[1:10,]), odd = '#eeeeee', even = 'white' )))

```



## Conditional formatting

```{r,echo=FALSE,message=FALSE,results='asis', warning=FALSE,comment=FALSE}
# the default cellProperties
myCellProps = cellProperties( padding = 5 )
MyFTable = FlexTable( data = mtcars[1:10,], header.text.props = textBold()
  , header.cell.props = myCellProps
  , body.cell.props= myCellProps )



# MyFTable[mtcars$hp < 100,] = chprop( myCellProps, background.color = '#666633')
```




## Using gridExtra

```{r,echo=FALSE,message=FALSE,results=FALSE, warning=FALSE,comment=FALSE}
library(gridExtra)

tbl = tableGrob(mtcars[1:10,]
                # , show.rownames = FALSE
                )

# Use grid.arrange to generate the table for your Markdown output:
grid.arrange(tbl)

```



## DT interface to DataTables library

```{r,echo=FALSE,message=FALSE,results=FALSE, warning=FALSE,comment=FALSE}
# https://rstudio.github.io/DT/

library(devtools)
# install_github("rstudio/DT")


library(DT)
# datatable(iris)


# datatable(head(iris), class = 'cell-border stripe')

```


## In conclusion



